% T I T L E   P A G E
% -------------------
% Last updated Feb. 25, 2014, by Stephen Carr, IST-Client Services, to add a title to the Author's Declaration page.

% The title page is counted as page `i' but we need to suppress the
% page number.  We also don't want any headers or footers.
\pagestyle{empty}
\pagenumbering{roman}

% The contents of the title page are specified in the "titlepage"
% environment.
\begin{titlepage}
        \begin{center}
        \vspace*{0.6cm}

        \Huge
        {\bf An Assistive Handwashing System with Emotional Intelligence }
        
        \vspace*{0.4cm}
        \LARGE
        Using Emotional Intelligence in Cognitive Intelligent Assistant Systems
        
        \vspace*{0.8cm}

        \normalsize
        by \\

        \vspace*{0.8cm}

        \Large
        Luyuan Lin \\

        \vspace*{2.8cm}

        \normalsize
        A thesis \\
        presented to the University of Waterloo \\ 
        in fulfillment of the \\
        thesis requirement for the degree of \\
        Master of Science \\
        in \\
        Artificial Intelligence \\

        \vspace*{1.6cm}

        Waterloo, Ontario, Canada, 2014 \\

        \vspace*{0.8cm}

        \copyright\ Luyuan Lin 2014 \\
        \end{center}
\end{titlepage}

% The rest of the front pages should contain no headers and be numbered using Roman numerals starting with `ii'
\pagestyle{plain}
\setcounter{page}{2}

\cleardoublepage % Ends the current page and causes all figures and tables that have so far appeared in the input to be printed.
% In a two-sided printing style, it also makes the next page a right-hand (odd-numbered) page, producing a blank page if necessary.
 


% D E C L A R A T I O N   P A G E
% -------------------------------
  % The following is the sample Delaration Page as provided by the GSO
  % December 13th, 2006.  It is designed for an electronic thesis.

\begin{center}\textbf{Author's Declaration}\end{center}

  \noindent
I hereby declare that I am the sole author of this thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

  \bigskip
  
  \noindent
I understand that my thesis may be made electronically available to the public.

\cleardoublepage
%\newpage

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}

Whether emotional intelligence (or, affective reasoning) is included can influence the effectiveness of a cognitive assistive system. This thesis presents a novel emotionally intelligent hand-washing assistant that aims to help older adults Alzheimer's disease complete hand-washing tasks more independently. The thesis reviews previous works in the development of cognitive assistants and in the study of emotional intelligence, and then designs a hand-washing system prototype that combines the two research streams. The difficulties in designing such a system, including probabilistic and decision-theoretic reasoning of the user's functional and emotional states, computer-vision based activity monitoring and affective recognition, and embodied prompting are discussed. Designing the hand-washing system as an integration of independent components, the thesis also discussed coordination between the components. The thesis implements the system in the end, and shows by preliminary tests in laboratory settings that the system implemented (1) runs in real-time from the perspective of the user group, (2) is able to provide a level of functional assistance, (3) produces system prompts that have encoded to some extent the emotional state of its user. The preliminary tests also indicated that an user with emotions with high potency levels (and high activity levels) is more likely to receive system prompts with low potency levels (and high activity levels).

This thesis is one of the exploratory works in the area of integrating emotional intelligence with cognitive intelligent assistive systems. It provides a solution to fitting emotional intelligence in a functional system, as well as points out directions for future improvements. The framework designed in this thesis is portable and extensible, and can be generalized to be used in other applications.

\vskip 1cm
\noindent\textbf{Keywords.} Affective reasoning, emotional intelligence, Affect Control Theory, BayesACT, handwashing assistive system, affect recognition, affect signal generation, cognitive intelligent assistant.

\cleardoublepage
%\newpage

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgements}\end{center}

I would like to take this opportunity to thank all the people who made this thesis possible.

Foremost, I would like to express my sincere gratitude to my supervisor Prof. Jesse Hoey for the continuous support to my master's study and research. His guidance helped me in all the time of research and writing of this thesis. 

I would like to thank the rest of my thesis committee: Prof. James Tung and Prof. Peter van Beek, for their insightful comments and questions.

I would also like to thank my friends and fellow labmates in the Health Informatics Group: Chengbo Li and Xiao Yang, and my boyfriend Enxun Wei, for the stimulating discussions, valueble advices in writing this thesis, and for all the fun we have had together. Additional thanks goes to my boyfriend who accompanied me for the sleepless nights writting this thesis.

Last but not the least, I would like to thank my parents, my little sister, and all my other friends, for their constant encouragement.

\cleardoublepage
%\newpage

% D E D I C A T I O N
% -------------------

%\begin{center}\textbf{Dedication}\end{center}
% e.g. This thesis is dedicated to the one I love.
%\cleardoublepage
%\newpage

% T A B L E   O F   C O N T E N T S
% ---------------------------------
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection
%\newpage

% L I S T   O F   T A B L E S
% ---------------------------
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page
%\newpage

% L I S T   O F   F I G U R E S
% -----------------------------
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page
%\newpage

% L I S T   O F   S Y M B O L S
% -----------------------------
% To include a Nomenclature section
% \addcontentsline{toc}{chapter}{\textbf{Nomenclature}}
% \renewcommand{\nomname}{Nomenclature}
% \printglossary
% \cleardoublepage
% \phantomsection % allows hyperref to link to the correct page
% \newpage

% Change page numbering back to Arabic numerals
\pagenumbering{arabic}

